### 2017-10-30 manuscript_gravitationalconstant.tex
corrections

\documentclass[prb,preprint]{revtex4-1} 
% The line above defines the type of LaTeX document.
% Note that AJP uses the same style as Phys. Rev. B (prb).

% The % character begins a comment, which continues to the end of the line.

\usepackage{amsmath}  % needed for \tfrac, \bmatrix, etc.
\usepackage{amsfonts} % needed for bold Greek, Fraktur, and blackboard bold
\usepackage{graphicx} % needed for figures
\usepackage{epstopdf}

\begin{document}

% Be sure to use the \title, \author, \affiliation, and \abstract macros
% to format your title page.  Don't use lower-level macros to  manually
% adjust the fonts and centering.

\title{A scenario for a Planckian standard kilogram}
% In a long title you can use \\ to force a line break at a certain location.

\author{Michael Ebersold}
\email{michael.ebersold@uzh.ch} 
% If there were a second author at the same address, we would put another 
% \author{} statement here.  Don't combine multiple authors in a single
% \author statement.
\author{Prasenjit Saha}
\email{psaha@physik.uzh.ch}
% Please provide a full mailing address here.
\affiliation{Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland}

% See the REVTeX documentation for more examples of author and affiliation lists.

\date{\today}

\begin{abstract}
The kilogram could, in principle, be redefined as an integer multiple
of the Planck mass (or 4594xxxx$\times\sqrt{\hbar c/G}$) with no
significant change from its current value.  The practical hurdle is
measuring $G$ three or four orders of magnitude more precisely than presently known,
so as to find the unknown digits xxxx.  We study a possible scenario
for doing so with an artificial two-body orbit.  Requirements for
suppressing perturbing effects (notably tidal forces and radiation
pressure), and obtaining the desired signal-to-noise are estimated.
\end{abstract}

\maketitle

\section{Introduction}

The reform of SI units to use quantum standards is an ongoing `news
story' in physics.  At the heart of the reform are two quantum
phenomena that appear on macroscopic scales: the Josephson effect
makes $h/e$ measurable as a voltage-to-frequency ratio, while the
quantum Hall effect makes $h/e^2$ a standard resistance (the
von~Klitzing constant).  These enable electrical units to be defined
in terms of $h$ and $e$.  A further ingenious instrument, known as a
watt balance, precisely compares electrical and mechanical power, thus
enabled mass to be standardized in terms of $h$.  For a simple
introduction to the new SI, we refer the reader to the article on the
``LEGO watt balance''[2015AmJPh..83..913C].\cite{Chao15}

The proposed redefinition of the kilogram is as an explicit value for
$h$ and $c$, but we can think of it as
\begin{equation}
\mathrm{kg} \equiv 1.3563925\ldots \times 10^{50} \, \mathrm{s}^{-1}
                   \times \frac{h}{c^2}
\end{equation}
with the exact value to be decided.  As we can see, the kilogram is
coupled to the definition of a second.  That is not a problem in
practice, as the second is standardized orders of magnitude better
than any other unit.  Nonetheless, it is interesting to consider a
logical further step of decoupling the kilogram from other
definitions.  This could be done by introducing the gravitational
constant, since $\sqrt{hc/G}$ has dimensions of mass.

Historically, there have been several proposed unit systems that
included $G$ in some way,\cite{Tomilin1999} but the idea of including
both gravity and quantum mechanics in a unit system comes from
Planck.\cite{Planck99}  In the process of working towards the
radiation law, Planck realized that $c,G,h$ and the Boltzmann constant
$k_B$ (in his notation $G\rightarrow f$, $h\rightarrow b$,
$h/k_B\rightarrow a$) naturally define units for mass, length, time,
and temperature.  (Planck remarked that such units would be meaningful
even in non-human and extraterrestrial cultures.)

A hypothetical Planckian SI unit system would have
\begin{equation} \label{eq:defs}
\begin{aligned}
\mathrm{kg} &= 4.594\ldots \times 10^7
               \left(\frac{\hbar c}{G}\right)^{\!1/2} \\
\mathrm{m}  &= 6.187\ldots \times 10^{34}
               \left(\frac{\hbar G}{c^3}\right)^{\!1/2} \\
\mathrm{s}  &= 1.854\ldots \times 10^{43}
               \left(\frac{\hbar G}{c^5}\right)^{\!1/2} \\
\mathrm{K}  &= 7.058\ldots \times 10^{-33}
               \left(\frac{\hbar c^5}{Gk_B^2}\right)^{\!1/2}
\end{aligned}
\end{equation}
with the numerical coefficients chosen so as to match the present
standards.  The square-rooted expressions are conventionally known as
the Planck mass, Planck length, and so on.  (They are actually
$\sqrt{2\pi}$ times Planck's definitions.)  Extension to electrical
units is straightforward, but we will not discuss it here.  The Planck
length, time, and temperature themselves are extreme scales, not
probed by any known physical process, and presumably the regime of
quantum gravity.  But extreme scales do not harm if they are only part
of some expression, as they are here.  The Planck mass is small but
microscopically measurable: $\sim20\mu\mathrm{g}$.  This is of the
same order as the uncertainty in the prototype kilogram itself.\cite{BIPM}  Hence
the kilogram could be redefined as an integer multiple of the Planck
mass, without significantly perturbing its value.

The difficulty with Eq.~\eqref{eq:defs} is that all the expressions
involve $G$.  Given the current uncertainties in $G$,\cite{CODATA16}
Eq.~\eqref{eq:defs} would allow units to be standardized only to four
significant digits, which is obviously not enough.  Improvement by
several orders of magnitude would be needed, to make the Planckian
kilogram practicable.  In this paper, we study a scenario which may
make such an improvement possible.  A Planckian second is not a prospect
with anything like present technology, because demands on the time
standard are more stringent still.\\

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{img/plotGmeas}
	\caption{Six of the most recent measurements of the gravitational constant $G$ with their uncertainties are plotted. The large disagreement is recognised immediately. The green lines represents the current recommended CODATA value and its uncertainty.}
	\label{fig:Gmeasurements}
\end{figure}

\section{A new approach to measure $G$}

In Fig.~\ref{fig:Gmeasurements} we have plotted some of the most recent measurements of the gravitational constant. Three of them have been carried out using a torsion balance [(BIPM-14)\cite{Quinn14}, (UCI-14)\cite{Newman14}, (HUST-09)\cite{Luo09}], (UZur-06)\cite{Schlamminger06} used a beam balance and (JILA-10)\cite{Parks10} measured the separation of two simple pendulums in a gravitational field. These measurements all tried to quantify the very weak gravitational force with a mechanical apparatus. The sixth measurement, (LENS-14)\cite{Rosi14}, was performed under another approach, using cold atom interferometry. Its uncertainty is a lot larger than the others, but it could become more competitive in the future. However, a brief glance on the plotted measurements shows the disagreement of current $G$-values and suggests a demand for new, potentially different techniques.\\

\subsection{Description of new measurement}
Newton's law of gravitation does not only explain why an apple falls down from a tree, it also explains orbits. If one has two bodies of masses $m_1$ and $m_2$, on which the only force acting is their respective gravitational attraction, they will orbit around their common barycenter in an ellipse described by Kepler's laws. The third law, given in Eq.~(\ref{eq:kepler3}), states that the square of the orbital period $P$ is directly proportional to the cube of the semimajor axis $a$. In this proportionality factor, the gravitational constant $G$ appears.
\begin{equation}\label{eq:kepler3}
P^2 = \frac{4 \, \pi^2}{G(m_1+m_2)}a^3
\end{equation}
If one is able to measure the position of the two masses very accurately and thus knows the orbit, it is possible to deduce the gravitational constant. 
So what we need is a freely falling two-body system. The only way to achieve this over some time, is to go to space. Thus, one has to design a spacecraft involving a cavity that contains the two masses and a position measurement apparatus. The cavity should shield the two-body system from environmental influences. The spacecraft is then put in space, preferably to a Lagrange point of the earth-sun system, in order to prevent large tidal forces.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\textwidth]{img/sketchnonumb.pdf}
	\caption{Sketch on how the two masses inside the cavity of a spacecraft orbit around each other.}
	\label{fig:sketch}
\end{figure}

To get a feeling what the dimensions of such an experiment could be, choose for example lead balls with a diameter of about 10 cm, their mass then is about 6 kg. If one then chooses a semimajor axis of 15 cm, the orbital period would be roughly 3.5 hours and the gravitational force acting between the two bodies is of order $10^{-8}$ N. In Fig.~\ref{fig:sketch} we show a sketch on how this would look like. Treating this system with Newtonian gravity is justified, because corrections from general relativity would be of order $(v^2/c^2)$ and $(GM/rc^2)$, which for our two-body orbit are both very small numbers.\\

For getting an improvement in the precision of the gravitational constant, the positions of the two balls have to be known very accurately. Best suitable would probably be a laser interferometry system, which can yield precisions down to some nanometers.\cite{Loughridge13}
The main difference to other techniques is that position measurements can be very accurate compared to directly measuring the gravitational force.
This new kind of measurement of the gravitational constant appears simple, but it brings various difficulties with it. It is practically impossible to have a clean system, where only the respective gravitational force is acting between the two bodies, so one has to find out what effects could possibly disturb the system and then either make them as small as possible or correct for them. \\

\subsection{Disturbing effects}

\textbf{Radiation pressure:} The envelope of the cavity needs to protect the two body-system from the constant bombardment of radiation and charged particles from the sun. Since the activity of the sun varies much, the spacecraft must be able to correct its position around the freely falling masses inside the cavity.\\
Furthermore, there is thermal radiation inside the cavity, that could disturb the orbit.
Using the Stefan-Boltzmann law, we can relate radiation pressure to the temperature inside the cavity. If we then specify to balls of radius $r$, we can express the force on the balls as 
\begin{equation}
F = \frac{8 \pi^6 r^2 {k_\mathrm{B}}^4 T^4}{45 c^3 h^3}.
\end{equation}
As we can see, this force is highly temperature dependent. In order to minimize this influence, one should deploy a temperature as small as possible in the cavity. In fact, if a cooling system based on liquid helium is installed, temperature can be brought down to about 4 K. Such a low temperature results in a force on the balls in the order of $10^{-16}$ N. Compared to the gravitational force, this is orders of magnitude less and thus we don't have to worry about effects of radiation pressure.\\

\textbf{Influence of the position measurement:} 
The photons of the laser that measures the position of the two balls exert an additional force. If a laser with a power of 1 mW is used and not more than five measurements every second are performed, the force on the balls is in the same order as the influence of thermal radiation and thus negligible too.\\

\textbf{Coulomb force:}
Since the electromagnetic force is orders of magnitude stronger than the gravitational force, the balls must be absolutely neutral. A possible source of charges are cosmic rays or the solar wind, so they have to be protected as good as possible. If we want that the Coulomb force between the balls is in the same order as the effect of thermal radiation and thus negligible, the maximal allowed charge on the balls is about $6 \cdot 10^{-14}\,\mathrm{C}$, which corresponds to roughly $4 \cdot 10^5$ elementary charges.\\

\textbf{Coriolis effect:}
\label{sec:coriolis}
The Coriolis force is a fictitious force which only acts on moving bodies in a rotating reference frame. If we imagine the spacecraft located stationary at a Lagrange point of the earth-sun system, it is situated in a rotating frame around the earth-sun barycenter. Since the balls are moving with respect to this frame, there will be a Coriolis acceleration
\begin{equation}
\label{eq:coriolis}
\ddot{\vec{r}}_\mathrm{c} = - 2 \; \vec{\omega} \times \dot{\vec{r}}.
\end{equation}
In this formula $\vec{\omega}$ is the angular velocity vector of the rotating frame and $\dot{\vec{r}}$ is the velocity vector in the two-body system.\\

\textbf{Centrifugal and tidal forces:}
\label{sec:tidal}
The free fall condition is not possible to achieve for both masses simultaneously since we are in a non-inertial, rotating frame and therefore have to account for centrifugal forces. The centrifugal acceleration $\ddot{\vec{r}}_{\mathrm{Z}}$ is proportional to the distance vector $\vec{r}$ and scales with $\omega^2$, where $\omega$ is the angular velocity of the sun-earth reference frame:
\begin{equation}
\label{eq:centrifugalacc}
\ddot{\vec{r}}_{\mathrm{Z}} = \vec{\omega} \times \left( \vec{\omega} \times \vec{r} \right) 
\end{equation}
On the other hand, there are tidal forces. Imagine if one mass is closer to the sun-earth barycenter, the gravitational force on it is slightly bigger than on the other mass. This results in a correction proportional to the distance too. One can approximate the tidal acceleration as
\begin{equation}
\label{eq:tidalacc}
\ddot{\vec{r}}_{\mathrm{\tau}} \approx \frac{2 \, G \, M_\odot}{R^3} \, \vec{r}.
\end{equation}
In this equation $M_\odot$ is the solar mass and $R$ is the radius of the orbit around the sun. With Kepler's third law one can rewrite the tidal acceleration as a function of the orbital period squared of the sun-earth reference frame. Thus it scales the same way as the centrifugal force. To estimate the magnitude of the centrifugal and tidal acceleration relative to the gravitational acceleration between the two balls, we find that their ratio is proportional to the orbital periods squared
\begin{equation}\label{a_ratio}
\frac{\ddot{r}_{\mathrm{\tau}}}{\ddot{r}_\mathrm{G}} \propto \frac{P^2}{{P_\odot}^2} \approx 10^{-7},
\end{equation}
here $P_\odot$ corresponds to one year, whereas $P$ is the orbital period of the two balls. So tidal and centrifugal effects are about $10^{-7}$ times smaller than the respective gravity of the two balls.\\

\textbf{Gravitational field:}
Much attention has to be paid to the effect of the gravitational field of the spacecraft itself and the measurement apparatus. This certainly can't be neglected, but with a clever design, its effects can be minimized. For example an almost spherically symmetric spacecraft would probably be a good choice. It should also be possible to calculate the gravitational field caused by the components of the spacecraft at the positions of the balls. Then one can correct the equation of motion by adding an extra term that is accounting for the gravitational field. \\
A source of an additional gravitational field is earth's moon and the other planets in the solar system. Fortunately, for this we don't need to correct if the whole system, consisting of the spacecraft and the two-body system, is moving through space together. Although, the spacecraft should be able to correct its position around the two balls to avoid a collision.\\

\section{Extracting $G$ from position measurement data}

In this section we are going to show how the gravitational constant can be recovered from position data of the two balls that the laser interferometry measurement would provide.
But first we have to find the orbital equation of motion for the two balls.\\

\subsection{Orbital equation of motion}
From Newtonian dynamics the equation of motion for a clean two-body system is given by 
\begin{equation}
\label{eq:eomeasy}
\ddot{\vec{r}} = - \frac{G M\, \vec{r}}{r^3}.
\end{equation}
Here $M$ denotes the sum of the masses $m_1$ and $m_2$, $G$ is the gravitational constant and $r$ is the relative distance between the two bodies. For simplicity we will abbreviate the product of $GM$ with $k$.
Since our two-body system is in a rotating frame, we have to add the term developed before for the Coriolis acceleration as well as a correction term proportional to the distance accounting for centrifugal and tidal forces. This last term can be written with a
symmetric $3 \times 3$ matrix $C$ as follows  
\begin{equation}
\label{eq:tidal}
\ddot{\vec{r}}_{\mathrm{C}} = C \cdot \vec{r}.
\end{equation}
or in terms of the individual components 
\begin{equation}
\label{eq:tidacom}
\left(\ddot{r}_\mathrm{C}\right)_i = \sum_{k=1}^{3} C_{ik} \: r_k.
\end{equation}
It can be thought of as a first order expansion and therefore also include other possible minor perturbations. The reference frame is chosen such that the barycenter of the two-body system is force-free.
So the full equation of motion for the two balls looks as follows:
\begin{equation}
\label{eq:eom}
\ddot{\vec{r}} = - \frac{k\,\vec{r}}{r^3} - 2 \; \vec{\omega} \times \dot{\vec{r}} + C \cdot \vec{r}
\end{equation}
These are actually three coupled equations of motion, one for each coordinate axis. Since all are second order differential equations, this leaves us with six initial conditions. In other words to determine the orbit, we need to give six starting values.\\
We choose our coordinate system such that the orbital plane of the two-body system is in the $x-y$ plane. $\vec{\omega}$ is a constant vector that points in the direction of the rotation axis of the earth-sun reference frame. We will fit for $\vec{\omega}$, so we don't have to know it exactly before, but it can be estimated well and thus provides good starting points for the fit. The same is valid for the six components of the matrix $C$.\\

Counting all unknown parameters in the equation of motion, we get 16. They correspond to 6 initial conditions, 6 components of the matrix $C$, 3 components of the vector $\vec{\omega}$ and the object of main interest, $k$,  which contains the gravitational constant ($k = GM$). Fitting for so many parameters is not an easy task, but it should be possible because the parameters are not completely unknown.\\

\subsection{Markov Chain Monte-Carlo fit}

Our problem has 16 unknown parameters, but the only one we are really interested in is $k$, which contains the gravitational constant. For the fit, we suggest to use a Markov Chain Monte-Carlo (MCMC) method because it does good work even if the parameter space is large. 
A Markov chain Monte-Carlo method is a Bayesian approach to data analysis. It uses the information provided by observed data about a set of parameters, to update a prior state of beliefs about the set of parameters to become a posterior state of beliefs about a set of parameters. We will not describe in detail how the MCMC is working, rather refer to an online tutorial from ESO PyCoffee\cite{esopycoffee} or an introduction paper from van Ravenzwaaij et al.\cite{vanRavenzwaaij16}

First we have to specify the prior probability distribution, so to say what values for the 16 parameters in the equation of motion actually make sense. Then, for all 16 parameters, a uniform distribution within these ranges and zero probability everywhere else is assumed as the prior probability distribution.
We briefly explain in what range the values of the parameters are estimated and display it in Table~\ref{tab:parameters}.
For the parameter $k$, which is the product of the gravitational constant $G$ and the total mass $M$, if we look at the current measured values for the gravitational constant in Fig.~\ref{fig:Gmeasurements}, it is reasonable to constrain its numerical value in SI units to values between $6.671 \cdot 10^{-11}$ and $6.676 \cdot 10^{-11}$. The masses of the two balls will not contribute to a larger uncertainty in $k$, we choose them to be 6 kg.

The vector $\vec{\omega}$ accounts for the rotation of the reference frame. If we say that the orbit of the balls lies in the same plane as the orbital plane of the earth, $\vec{\omega}$ will point in $z$ - direction with $x$ and $y$ components small. Thus $\omega_z$ is associated with the period $P_\odot$ of one year through $2\pi/P_\odot$. Because the rotation axis are conceivably not exactly aligned, we add some uncertainty. As mentioned before, the $x$ and $y$ components are small, so to say roughly in the range of the uncertainty in $w_z$.
The main contribution to the tidal term goes with $\omega^2$ and lies in the orbital plane. We are free to choose the $x$ and $y$ axis, so we choose it such that it is in $x$-direction. The other five components then should be smaller, but we give some extra range to account for possible other perturbations.
The initial values for the positions can be guessed very accurately, since they are directly measured precisely. The initial velocity can be calculated easily and matched to the initial positions, therefore also yielding narrow parameter ranges.

\begin{table}[h]
	\centering
	\caption{This table shows the ranges in which the parameters numerical values in SI units are estimated to make sense and the value which was chosen to generate data.}
	\begin{ruledtabular}
		\begin{tabular}{l c c}
			% The codes above determine the horizontal alignment in each column.
			% Options are l (left), r (right), c (centered), and p (paragraph).
			% The p option allows an entry to be broken into multiple lines, and
			% therefore requires a width specification, in this case 5 centimeters.
			parameter & range & chosen value \\
			\hline	% horizontal line to separate headings from data
			$k$ & $(8.0052$--$8.0112) \cdot 10^{-10}$ & $8.0088 \cdot 10^{-10}$ \\
			$\omega_x$ & $\pm 1.0 \cdot 10^{-8}$ & $0.02 \cdot 10^{-8}$ \\
			$\omega_y$ & $\pm 1.0 \cdot 10^{-8}$ & $-0.03 \cdot 10^{-8}$ \\
			$\omega_z$ & $(1.95$--$2.05) \cdot 10^{-7}$ & $1.996 \cdot 10^{-7}$ \\
			$C_{xx}$ & $0$--$5.0 \cdot 10^{-14}$ & $3.8 \cdot 10^{-14}$ \\
			$C_{yy}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{zz}$ & $\pm 10^{-14}$ & $-0.2 \cdot 10^{-14}$ \\
			$C_{xy} = C_{yx}$ & $\pm 10^{-14}$ & $0.3 \cdot 10^{-14}$ \\
			$C_{yz} = C_{zy}$ & $\pm 10^{-14}$ & $0.1 \cdot 10^{-14}$ \\
			$C_{xz} = C_{zx}$ & $\pm 10^{-14}$ & $-0.4 \cdot 10^{-14}$ \\
			$x_0$ & $0.15 \pm 10^{-7}$ & $0.15$ \\
			$y_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$z_0$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{x_0}$ & $\pm 10^{-7}$ & $0.0$ \\ 
			$v_{y_0}$ & ($7.30$--$7.31) \cdot 10^{-7}$ & $7.30698296 \cdot 10^{-5}$ \\
			$v_{z_0}$ & $\pm 10^{-7}$ & $0.0$ \\
		\end{tabular}
	\end{ruledtabular}
	\label{tab:parameters}
\end{table}

In the next step, we set the parameters to a fixed ``true" value within the range estimated. We have to keep in mind that $k$, $x_0$ and $v_{y_0}$ are linked together and there must not be contradictions in the true values. The chosen values are displayed in the right column in Table~\ref{tab:parameters}.
To simulate data we would obtain from the position measurements, we integrate the equations of motion. The time for one orbit without perturbations can be estimated using Kepler's third law (\ref{eq:kepler3}). For the value of $k$ assumed above and a semi-major axis of $0.15 \, \mathrm{m}$, this results in an orbital period of roughly $12900 \, \mathrm{seconds}$ or $3.5 \, \mathrm{hours}$. Here we integrate two orbits in $100000$ steps, that would correspond to about $4$ measurements per second, which should be feasible. To the generated data, we add normal distributed noise with a standard deviation of $10 \, \mathrm{nm}$. This uncertainty stays for a systematic error in the laser interferometry measurement and also random statistic noise. Its value is just an estimation, it would depend highly on the measurement system that has been used.

\begin{figure}[h]
	\centering
	\includegraphics[width=1.0\textwidth]{img/4cornerfit.png}
	\caption{Corner plot of the posterior for 4 parameters: $x_0$, $v_{y_0}$, $\omega_z$ and $k$. Indicated with blue lines are the values we calculated the data from. The dashed lines in the histograms along the diagonal represent the median and the 0.16 and 0.84 percentiles respectively.}
	\label{fig:plotCorner4}
\end{figure}

A useful way to look at the samples of the posterior returned by the MCMC is doing a corner plot using a Python module named \textit{corner}\cite{corner}. The corner plot shows all the one and two dimensional projections of the posterior probability distributions and demonstrates quickly the covariances between parameters. Along the diagonal, the marginalized distribution for each parameter independently is shown.
In Fig.~\ref{fig:plotCorner4} an excerpt of the full corner plot is shown.

To arrive at a best value with uncertainty for the gravitational constant, we take the mean and standard deviation of the marginalized distribution of the parameter $k$ and divide by the total mass of the balls. In this example the relative uncertainty of $k$ is roughly $10^{-7}$, so if the masses are measured to about the same precision, we can write for the gravitational constant $G$:
\begin{equation}\label{Gfit}
G = \left(6.674005 \pm 0.000001 \right) \cdot 10^{-11}  \frac{\mathrm{m}^3}{\mathrm{kg \, s}^2}
\end{equation}
This corresponds to a relative uncertainty of $1.5 \cdot 10^{-7}$ or 15 parts in a hundred million.

\section{Conclusion}
We have proposed a new way to measure the gravitational constant $G$ and estimated on what order of magnitude different disturbing effects occur and if one has to correct for them. We then simulated measurement data and showed how to extract the gravitational constant. Based on our estimations, it should be possible to measure the gravitational constant to about $10^{-7}$ relative uncertainty. This is a about a factor of 300 better than today's recommended CODATA value.

If we are a bit optimistic, a further increase in precision is not out of reach, so that about the same relative uncertainty as for Planck's constant $h$ ($1.2 \cdot 10^{-8}$)\cite{CODATA16} could be achieved. This then would allow to redefine the kilogram as an integer multiple of the Planck mass.\\

\appendix*   % Omit the * if there's more than one appendix.

\section{Uninteresting stuff}

%Appendices are for material that is needed for completeness but
%not sufficiently interesting to include in the main body of the paper.  Most
%articles don't need any appendices, but feel free to use them when
%appropriate.  This sample article needs an appendix only to illustrate how 
%to create an appendix.

\begin{thebibliography}{99}

\bibitem{Chao15} L. S. Chao, S. Schlamminger, D. B. Newell, J. R. Pratt, F. Seifert, X. Zhang, G. Sineriz, M. Liu and D. Haddad, ``A LEGO Watt balance: An apparatus to determine a mass based on the new SI", \textit{American Journal of Physics}, 83:11:913--922 (2015).

\bibitem{Tomilin1999} D. K. Tomilin, ``Natural Systems of Units. To the Centenary Anniversary of the Planck System,'' 
Proceedings of the XXII Workshop on high Energy Physics and field theory, 290 (1999).

\bibitem{Planck99} M. Planck, ``\"{U}ber irreversible {S}trahlungsvorg\"{a}nge,'' 
Sitzungsberichte der K\"{o}niglich Preussischen Akademie der Wissenschaften zu Berlin, 479--480 (1899).

\bibitem{BIPM} Bureau International des Poids et Mesures, Calibration and measurement services: Mass, \url{<https://www.bipm.org/utils/common/pdf/cms_m.pdf>}, (2015).

\bibitem{CODATA16} P. J. Mohr, D. B. Newell and B. N. Taylor, ``CODATA recommended values of the fundamental physical constants: 2014*, \textit{Reviews of Modern Physics}, 88(3):035009 (2016).

\bibitem{Quinn14} T. Quinn, C. Speake, H. Parks and R. Davis, ``The BIPM measurements of the Newtonian constant of gravitation $G$'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (2014).

\bibitem{Newman14} R. Newman, M. Bantel, E. Berg and W. Cross, ``A measurement of $G$ with a cryogenic torsion pendulum'', \textit{Philosophical Transactions of the Royal Society of London A: Mathematical, Physical and Engineering Sciences}, 372(2026) (2014).

\bibitem{Luo09} J. Luo, W. Liu, L. Tu, C. Shao, L. Liu, S. Yang, Q. Li and Y. Zhang, ``Determination of the Newtonian gravitational constant $G$ with time-of-swing method'', \textit{Phys. Rev. Lett.}, 102:240801 (2009).

\bibitem{Schlamminger06} S. Schlamminger, E. Holzschuh, W. Kündig, F. Nolting, R. E. Pixley, J. Schnurr and U. Straumann, ``Measurement of Newton's gravitational constant", \textit{Phys. Rev. D}, 74:082001 (2009).

\bibitem{Parks10} H. V. Parks and J. E. Faller, ``Simple pendulum determination of the gravitational constant", \textit{Phys. Rev. Lett.}, 105:110801 (2010).

\bibitem{Rosi14} G. Rosi, F. Sorrentino, L. Cacciapuoti, M. Prevedelli and G. M. Tino, ``Precision measurement of the Newtonian gravitational constant using cold atoms", \textit{Nature}, 510:518--521 (2014).

\bibitem{Loughridge13} R. Loughridge and D. Y. Abramovitch, ``A tutorial on laser interferometry for precision measurements,'' 
American Control Conference, 3686--3703 (2013).

\bibitem{esopycoffee} Fast and easy-to-implement Markov Chain Monte Carlo with the emcee package, ESO PyCoffee (2014),
\url{<http://eso-python.github.io/ESOPythonTutorials/ESOPythonDemoDay8_MCMC_with_emcee.html>}.

\bibitem{vanRavenzwaaij16} D. van Ravenzwaji, P. Cassey and S. D. Brown, ``A simple introduction to Markov Chain Monte-Carlo sampling,'' 
Psychonomic Bulletin {\&} Review, 1--12 (2016).

\bibitem{corner} D. Foreman-Mackey, ``corner.py: Scatterplot matrices in Python,'' 
The Journal of Open Source Software (10), (2016).

\end{thebibliography}

\end{document}