### 2020-04-07 spg/main.tex
SPG article

\documentclass[aps,prb,12pt]{revtex4-1}

\def\unit#1{\;{\rm#1}}

\begin{document}

\title{Planck and the Extra-terrestrials\\
       or what the new SI means for astronomy}

\author{Prasenjit Saha} \email{psaha@physik.uzh.ch}
\affiliation{Physik-Institut, University of Zurich\\
  Winterthurerstr.~190, 8057~Zurich, Switzerland}

\begin{abstract}
New freedom.
\end{abstract}

\maketitle

In 1899 Planck \cite{Planck1899}.
Let $m=\sqrt{hc/G}$.
$f \rightarrow G$, $b\rightarrow h$, $a\rightarrow h/k$
Comes to $55\rm\,\mu g$.

Not the first \cite{Tomilin1999} but most prescient.  SI
\cite{Jeckelmann_2018}.  Leaves out $G$.  Also candela, sievert.

Astros use Planckian units occasionally (with $\hbar$) but generally
not.  Stories of cross-sections in gallons/Mpc.\footnote{A gallon/Mpc
  is about a kilogram, but the exact value depends on which gallon.}
Why?

Solar mass in kg would be bad.  SI in astro has been a non-starter.
But spacecraft dynamics uses $GM$ in SI.

New SI new freedoms.  eV not a new unit.  Density as
$\!\unit{eV}\unit{m}^{-3}$ is fine.  Anarchy?  No, because conversions
only through physical constants.

For astros, don't need to express solar mass in kg.  But we'll get to
that.

First Light-second: au and pc.  Again, not a new unit.  Round like Pfund.

Optical magnitude.  Photon flux on logarithmic wavelength scale.
Rounds nicely.  eNot to be confused with seismological magnitudes.

Gravity-second $GM/c^3$
(pulsars).
Strange formulations.
Dimensionless velocity ok.  But density $\rho$ as $s^{-2}$ seems
strange.  But $(G\rho)^{-1/2}$ is a time scale.  And gravity-second
per second is a power.  Basically, avoid explicit $G$ in pure
gravitational phenomena.  Explicit $G$ only when gravity and non-grav
processes both appear.

In the 120 years since Planck speculated about it, we have yet to make
contact with non-human or extra-terrestrial cultures, and ask them if
they indeed share Planck's views on units.  What {\em has} happened,
however, is that physical constants have turned out to have meanings
not even Planck in 1899 had imagined.  As we all know, the following
year Planck himself discovered that $h$ (formerly known as $b$) was
about quantization.  A few years later the formal relation between
mass and energy through $c^2$ turned out to be real.  Starting in the
1920s $(\hbox{Planck mass})^{-2}\times(\hbox{proton mass})^3$ was
discovered to be the mass scale of stars.  Later in the 20th century,
$2e/h$ and $h/e^2$ turned out to be macroscopically observable as
Josephson's constant and von~Klitzing's constant.  And just in the
last few years, $c^5/G$ has been observed to be the power-output scale
for merging black holes.  We can only wonder what more surprises await
us.

\def\aj{AJ}
\def\apjl{ApJL}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\pasj{PASJ}
\def\jgr{JGR}

\bibliographystyle{apsrev4-1}
\bibliography{refs.bib}

\end{document}